\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}

### 2022-04-19 sections/methods.tex
Change: Update the discussion section
Change: Move tables to the end of the manuscript

  \subsection*{16S rRNA sequencing Datasets}

@@

  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
  The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
  Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
  It also allows for automatic parallelization of all possible processes, both within and across samples.
  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
  % In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
  The main components of the pipeline are detailed in the subsequent sections.

@@

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
      \hline
      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
      \hline
      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Chimera checking & Remove bimera & \cite{Callahan2016} \\
      \hline
    \end{tabular}
    \caption{Tools used in the DC module}
    \label{tab:dc_tools}
  \end{table}

@@

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
      \hline
      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
      \hline
      Database & Greengenes & \cite{DeSantis2006} \\
      Database & SILVA & \cite{Quast2012} \\
      Database & NCBI RefSeq & \cite{Sayers2009} \\
      \hline
    \end{tabular}
    \caption{Tools used in the TA module}
    \label{tab:ta_tools}
  \end{table}

  \subsection*{OTU and ESV Processing (OP)}
  \vspace{-5mm}
  This module deals with normalization, filtering, forking, grouping and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
  Normalization of the count matrix involves converting the count matrix of read counts into a count matrix containing relative abundances.
  The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
  Although the pipeline supports rarefaction, these modules are turned off by default and the analyses reported in the paper do not normalize or rareify the counts matrices.
  However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
  By default, samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance of the 1\% and prevalence is less than 5\%.
  The forking operation splits the count matrix into multiple matrices based on sample metadata column.
  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy.
  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in analyses peformed in the study these are handled by the network inference algorithm).
  All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.

@@

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
      \hline
      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
      \hline
      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
      \hline
      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
      Correlation & Pearson & - \\
      Correlation & Spearman & - \\
      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
      \hline
    \end{tabular}
    \caption{Tools used in the NI module}
    \label{tab:ni_tools}
  \end{table}

@@

  % TODO: Should we remove CC from the LM model?
  In particular, we infer parameters $\alpha_i$ such that:
   \begin{equation*}
       C_j = \sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
             \sum_{k=1}^2 \left( \alpha^{CC(k)}_j.\delta^{CC(k)}_j \right) +
             \sum_{k=1}^3 \left( \alpha^{TA(k)}_j.\delta^{TA(k)}_j \right) +
             \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) +
             \sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
             \epsilon_j
   \end{equation*}

@@

>>>>

  \subsection*{16S rRNA sequencing datasets}

@@

  % TODO: Write about the JSON network format (supplementary)
  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
  The \ac{micone} Python API several biom file and network related function and methods, enabling detailed comparison of counts tables and inferred networks if desired.
  The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
  Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
  It also allows for automatic parallelization of all possible processes, both within and across samples.
  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
  % In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
  The main components of the pipeline are detailed in the subsequent sections.

@@

@@

  % TODO : Add information about filtering when OP=off
  \subsection*{OTU and ESV Processing (OP)}
  \vspace{-5mm}
  This module deals with normalization, filtering, forking, grouping and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
  Normalization of the count matrix involves converting the count matrix of read counts into a count matrix containing relative abundances.
  The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
  Although the pipeline supports rarefaction, these modules are turned off by default and the analyses reported in the paper do not normalize or rareify the counts matrices.
  However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
  By default, samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance of the 1\% and prevalence is less than 5\%.
  The forking operation splits the count matrix into multiple matrices based on sample metadata column.
  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy.
  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in analyses peformed in the study these are handled by the network inference algorithm).
  All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.

@@

@@

  % TODO: Should we remove CC from the LM model?
  In particular, we infer parameters $\alpha_i$ such that:
   \begin{equation*}
       C_j = \sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
             \sum_{k=1}^2 \left( \alpha^{CC(k)}_j.\delta^{CC(k)}_j \right) +
             \sum_{k=1}^3 \left( \alpha^{TA(k)}_j.\delta^{TA(k)}_j \right) +
             \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) + \\
             \sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
             \epsilon_j
   \end{equation*}

@@

  \subsection*{Statistical analyses}
  \vspace{-5mm}

  \subsubsection*{DC step}
  % Write about UniFrac

  \subsubsection*{TA step}
  % Write about Braycurtis

  \subsubsection*{NI step}
  % Write about precision and sensitivity equations used
  % L1 distance equations used

### 2022-04-19 sections/tables_figures.tex
Change: Figure 6 <-> Figure 7
Change: Move tables to the end of the manuscript

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{figure6.pdf}
    \caption{
      The precision and sensitivity of the inferred networks on the synthetic data.
    }
    \label{fig:figure6}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[h]
    \centering
    \includegraphics[width=17cm]{figure7.pdf}
    \caption{
      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
    }
    \label{fig:figure7}
  \end{figure}

>>>>

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
      \hline
      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
      \hline
      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
      Chimera checking & Remove bimera & \cite{Callahan2016} \\
      \hline
    \end{tabular}
    \caption{Tools used in the DC module}
    \label{tab:dc_tools}
  \end{table}

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
      \hline
      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
      \hline
      Database & Greengenes & \cite{DeSantis2006} \\
      Database & SILVA & \cite{Quast2012} \\
      Database & NCBI RefSeq & \cite{Sayers2009} \\
      \hline
    \end{tabular}
    \caption{Tools used in the TA module}
    \label{tab:ta_tools}
  \end{table}

  \begin{table}[h]
    \centering
    \small
    \begin{tabular}{|c|c|c|}
      \hline
      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
      \hline
      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
      \hline
      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
      \hline
      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
      Correlation & Pearson & - \\
      Correlation & Spearman & - \\
      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
      \hline
    \end{tabular}
    \caption{Tools used in the NI module}
    \label{tab:ni_tools}
  \end{table}

@@

  \begin{figure}[h]
    \centering
    \includegraphics[width=1.0\linewidth]{figure6.pdf}
    \caption{
      The precision and sensitivity of the inferred networks on the synthetic data.
    }
    \label{fig:figure6}
  \end{figure}
  \FloatBarrier
  \newpage

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{figure7.pdf}
    \caption{
      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
    }
    \label{fig:figure7}
  \end{figure}
